Lesson 3
========================================================

***

### What to Do First?
Notes:
This lesson explores analysis with one variable in a data set.  How does one varialbe change in a data set?  How to represent these changes in a graqph?  Other questons answered.
***

### Pseudo-Facebook User Data
Notes:
other way to load data from the file, use read.delim() function.  This function assumes that the tab character is the seperator between values and the period is the decimal character.
```{r Pseudo-Facebook User Data}
getwd()
list.files()

facebookData <- read.csv('pseudo_facebook.tsv', sep='\t')

facebookData2 <- read.delim('pseudo_facebook.tsv')

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}


# list the names of variables
names(facebookData2)
# now plot it, make sure the library() command i run first...
library(ggplot2)
qplot(x = dob_day, data = facebookData2) +
  scale_x_discrete(breaks=1:31)

```

***

#### What are some things that you notice about this histogram?
Response:
There are a lot of birthdaysa on the first of the month.  This seems very unusual.  Maybe the default for one's birthday is the first of the month selected?  Could be...

Fewer people born on the 31th...expected not all months have 31 days.
***

### Moira's Investigation
Notes:
outliers can be extreme cases.  Define extreme in context of rest of data.
Outlier to one data swet can be normal for another.

histogram axis can should be adjusted to show variations in data
  good to remove outliers first

Outliers, what can they be??:
  1. good data bout an exterme case
  2. bad data azbout an extreme case
  3. bad data about a non-extreme case
  
What is bad data?
  The system under study does not allow values that high/low?
  The system under study derives this data from other data, and that bada is bad.
  

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:
faceting means seperating the data using another variable that is not the variable under investigation,

Here is the problem faceting solves:
You want to do split up your data by one or more variables and plot the subsets of data together.

ncol = 3 means the number of clolumns 
```{r Faceting}
qplot(x = dob_day, data = facebookData2) +
  scale_x_discrete(breaks=1:31) +
  facet_wrap(~dob_month, ncol=3)

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:
Notes:
outliers can be extreme cases.  Define extreme in context of rest of data.
Outlier to one data swet can be normal for another.

histogram axis can should be adjusted to show variations in data
  good to remove outliers first

Outliers, what can they be??:
  1. good data bout an exterme case
  2. bad data azbout an extreme case
  3. bad data about a non-extreme case
  

***

### Moira's Outlier
Notes:
What is bad data?
  The system under study does not allow values that high/low?
  The system under study derives this data from other data, and that bada is bad.
#### Which case do you think applies to Moira’s outlier?
Response:
Bad data about an exterme case.  The system under study does not allow people to have more than 5,000 friends.

***

### Friend Count
Notes:
friend is represented by int
A lot of them...

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
names(facebookData2)
# now plot it, make sure the library() command i run first...
library(ggplot2)
qplot(x = friend_count, data = facebookData2) +
  scale_x_discrete(breaks=1:5000)
```

#### How is this plot similar to Moira's first plot?
Response:
The histogram is so heavily weighted to ward the low end, it is hard to distinguish individual values.  All mashed together...not clear.

An example of long-tail data.


***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = facebookData2,xlim=c(0,1000))

# another way, same histogram as before
qplot(x = friend_count, data = facebookData2,xlim=c(0,1000))
    scale_x_discrete(breaks=1:100)

# third way, this uses layers, after the + sign
qplot(x = friend_count, data = facebookData2) +
  scale_x_continuous(limits = c(0, 1000))
```

### Exploring with Bin Width
Notes:
What is bin width?  Threa the x-axix as a series of bins, each values on that axis contains items, guesses, for that values of x.  

Each value on the x-asix is a bucket containning positives, values, items.  The y-axis is how many items are in each x-axis bucket
***

### Adjusting the Bin Width
Notes:
Take a look at scale_x_continuous the parameters are:
  1. limits -> what is the range of allowed x values?
  2. breaks -> I am going from 0 to 1000, stepping every 50 units
  
when writing R code, the indenting is very important, like Python.

Also remember the + signs, and the tilde
### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = facebookData, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  # this is my added coded
  facet_wrap(~gender, ncol=1) 





```

***

### Omitting NA Values
Notes:
Two ways, na wrapper and by subset
  1. by subset -> data = subset(pf, !is.na(gender))
        subset(dataset, condition)
  2. by na.omit(pf) wrapper -> this omitts data with any N/A value in the record
      all or nothing!!!!
  
```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:
new function 'by'
by(variable in question, the categorial variable, and the data in question)

```{r Statistics \'by\' Gender}

```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!